---
title: "Mesocosm Light Test Data"
author: "Ruth"
date: "2023-06-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ruth is great at coding

Data loading and cleaning.

```{r}
library(readr)
AllData <- read_csv("AllData.csv")
#install.packages("ggplot2")
library("ggplot2")
#install.packages("tidyverse")
library("tidyverse")
#install.packages("lubridate")
library("lubridate")




AllData<- rename(AllData, Date = "Date Time, GMT-03:00", Temperature = "Temp, °C (LGR S/N: 1058708, SEN S/N: 1058708)"  , PAR = "PAR,  µmol photon m^-2 s^-1")


AllData$Date <- mdy_hms(AllData$Date)

AllData<- AllData %>%
  filter(Date < "2023-06-28 03:00:00")

```

PAR Plot 1 coding of all data. Plot 1 compares the PAR data from all 3 tanks and the wall.

```{r}
ggplot(AllData, aes(x=Date , y= PAR, color=Position)) + 
  geom_line() + 
  facet_wrap(vars(Tank)) + 
  ylab("PAR (µmol photon m^-2 s^-1)") +
  labs (title = "Underwater PAR Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "All Data", caption = "This graph compares the amount of light that 3 tanks recieved on the rim and underwater and compares that to the direct sunlight (Wall).")
```

New version of PAR Plot 1 not including the Rim and Wall Data.

```{r}

TankData<- AllData %>%
filter(Position != c('Rim')) %>%
  filter(Position != c('Wall')) 

ggplot(TankData, aes(x=Date , y= PAR, color=Position)) + 
  geom_line() + 
  facet_wrap(vars(Tank)) +
   ylab("PAR (µmol photon m^-2 s^-1)") +
  labs (title = "Underwater PAR Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "Only Data Collected Underwater Inside the Tanks", caption = "This graph compares the distribution of light underwater in each tank that was measured.")


```

New version of PAR Plot 2 with the Position and Tank variable switched around.

```{r}
ggplot(TankData, aes(x=Date , y= PAR, color=Tank)) + 
  geom_line() + 
  facet_wrap(vars(Position)) + 
  ylab("PAR (µmol photon m^-2 s^-1)") +
  labs (title = "Underwater PAR Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "Compares the Light Measurements in Different Areas Underwater Inside the Tanks", caption = "This graph compares the distribution of light throughout the bottom surface of each tank that was measured.")
```

New Temp Plot comparing the Temp data from all 3 tanks and the wall.

```{r}
ggplot(AllData, aes(x=Date , y= Temperature, color=Position)) + 
  geom_line() + 
  facet_wrap(vars(Tank)) + 
  labs (title = "Underwater Temperature Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "All Data", caption = "This graph compares the temperature that 3 tanks experienced on the rim and underwater and compares that to the direct sunlight (Wall).")
```

New verision of Temp Plot 1 excluding rim and wall data.

```{r}
ggplot(TankData, aes(x=Date , y= Temperature, color=Position)) + 
  geom_line() + 
  facet_wrap(vars(Tank)) +
  labs (title = "Underwater Temperature Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "Only Data Collected Underwater Inside the Tanks", caption = "This graph compares the distribution of temperature experienced underwater in each tank that was measured.")
```

New version of Temp Plot 2 with Position and Tank variable switched around.

```{r}
ggplot(TankData, aes(x=Date , y= Temperature, color=Tank)) + 
  geom_line() + 
  facet_wrap(vars(Position)) + 
  labs (title = "Underwater Temperature Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "Compares the Temperature Measurements in Different Areas Underwater Inside the Tanks", caption = "This graph compares the distribution of temperature experienced throughout the bottom surface of each tank that was measured.")
```
New Plot showing final calculated daily sunlight
```{r}
fcds <- read_csv("fcds.csv")

fcds <- rename(fcds, PAR = "PAR, mol photon m^-2 day^-1")

fcds_tank<- fcds %>%
filter(Position != c('Rim')) %>%
  filter(Position != c('Wall')) 


ggplot(data=fcds_tank, aes(x= Position, y= PAR, fill= Tank)) + 
  geom_bar(stat="identity") +
  facet_wrap(vars(Tank)) +
  ylab("PAR (mol photon m^-2 day^-1)") +
 labs (title = "Underwater PAR Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "Final Calculated Daily Sunlight", caption = "These three barplots compare the amount of light in mol photon m^-2 day^-1 that different locations inside the tanks received.")

#facet_wrap(vars(Tank)) + "PAR, mol photon m^-2 day^-1"
```
Mean and Standard Deviation of Final Calculated Daily Sunlight.
```{r}
Summary_Stats <- fcds_tank %>%
 group_by(Tank) %>%
 summarize (mean_PAR=mean(PAR, na.rm=TRUE), stdev_PAR=sd(PAR, na.rm=TRUE))

# Define the top and bottom of the errorbars
limits <- aes(ymax = mean_PAR + stdev_PAR, ymin=mean_PAR - stdev_PAR)

```

New Plot showing mean and standard deviation.
```{r}
ggplot(data=Summary_Stats, aes(x= Tank, y= mean_PAR, fill=Tank)) + 
  geom_bar(aes(color= Tank),stat="identity") + geom_errorbar(limits, stat= "identity", width=0.1) +
    ylab("Mean PAR (mol photon m^-2 day^-1)") +
 labs (title = "Underwater PAR Measurements Over a 48 Hour Time Period in the Mesocosm", subtitle = "Final Calculated Daily Sunlight - Mean and Standard Deviation", caption = "This bar graph compares the average amount of light in mol photon m^-2 day^-1 that each tank received.")
```

